Notes about the Minimal Dimensional Method to Classify Topological Phases (preliminary)

### 2017-07-25 Notes of various papers/Chiu Minimal Dirac Hamiltonian Method/chapters/chap-2.tex
[minor] Correct minor grammatical mistakes.

We begin our discussion with a background before this classification.  The
classification of topological phases with different non-unitary symmetries in
non-interacting picture, are in essence, the classification of $N\times N$
matrices by its response to three discrete symmetries: time reversal symmetry
($T$), charge conjugation/particle hole symmetry ($C$), and chiral/sub-lattice
symmetry ($S$). As for those unitary symmetries, one can in principle, block
diagonalize the Hamiltonian
$H=\mathrm{diag}(H^{\lambda_1},H^{\lambda_2},\cdots)$, such that each block
$H^{\lambda_i}$ is labeled by irreducible representations $\lambda_i$, and it
has no memory of the unitary symmetries. Therefore, our classification is
applied to those blocks and still holds.

It will be find that, those $T,C$ are the only two meaningful non-unitary
symmetries of the Single-particle system, and there are in total 10 different
ways in which the Single-particle Hamiltonian could respect the three symmetries
$T,C,S$. Now we explain in detail, about the reason why we care about, and only
care about these three symmetries $T,C,$ and $S$.

@@

But the square of $S$ is tricky. Since $S=TC=u_T u_C^*$, and each unitary matrix
$u_T$ and $u_C$ has a phase freedom (as they acts on creation/annihilation
operators), we can always pick a phase such that $S^2$ is some phase we want.
Sometimes we pick $S^2=1$. Sometime we want $\{T,C\}=0$, and pick some specific
phase for $S^2$.\footnote{This is a bit troublesome. First we denote
	$T^2=\varepsilon_T$, $C^2=\varepsilon_C$, then it can be found that
	$(u_T)^t=\varepsilon_T u_T$, $(u_C)^t = \varepsilon_C u_C$. Then, with
	$S=u_Tu_C^*$, we have $S^\dagger=\varepsilon_C\varepsilon_Tu_Cu_T^* =
	\varepsilon_C\varepsilon_T CT$, or $CT= \varepsilon_C\varepsilon_TS^\dagger$.
	Since $TC+CT=0$, we have $S+\varepsilon_C\varepsilon_T S^\dagger=0$, or
	$S^2=-\varepsilon_C\varepsilon_T$.}

@@

Since we are classifying in the solids, it is natural to expect translational
symmetry to present. Also, the addition of translational symmetry not only
produces the same classification, but also make this classification more
convenient to be done.

With translational symmetry added, the classification can be made the will yield
the same result.\footnote{This is still a mystery to me which this is true. The
	exact reason is related to some isomorphism between certain homotopy groups.
	Interested reader could see Ludwig's lecture
	\href{http://boulderschool.yale.edu/2016/boulder-school-2016-lecture-notes}{here}
	for a clue.}

>>>>

We begin our discussion with a background for this classification.  The
classification of topological phases with different non-unitary symmetries in
non-interacting picture, are in essence, the classification of $N\times N$
matrices by its response to three discrete symmetries: time reversal symmetry
($T$), charge conjugation/particle hole symmetry ($C$), and chiral/sub-lattice
symmetry ($S$). As for those unitary symmetries, one can in principle, block
diagonalize the Hamiltonian
$H=\mathrm{diag}(H^{\lambda_1},H^{\lambda_2},\cdots)$, such that each block
$H^{\lambda_i}$ is labeled by irreducible representations $\lambda_i$, and it
has no memory of the unitary symmetries. Then, our classification is
applied to those blocks and still holds.

It will be found that, those $T,C$ are the only two meaningful non-unitary
symmetries of the Single-particle system, and there are in total 10 different
ways in which the Single-particle Hamiltonian could respect the three symmetries
$T,C$, and $S=TC$. Now we explain in detail, about the reason why we care about, and only
care about these three symmetries $T,C,$ and $S$.

@@

But the square of $S$ is tricky. Since $S=TC=u_T u_C^*$, and each unitary matrix
$u_T$ and $u_C$ has a phase freedom (as they acts on creation/annihilation
operators), we can always pick a phase such that $S^2$ is some phase we want.
Sometimes we pick $S^2=1$. Sometime we want $\{T,C\}=0$, and pick another specific
phase for $S^2$.\footnote{This is a bit troublesome. First we denote
	$T^2=\varepsilon_T$, $C^2=\varepsilon_C$, then it can be found that
	$(u_T)^t=\varepsilon_T u_T$, $(u_C)^t = \varepsilon_C u_C$. Then, with
	$S=u_Tu_C^*$, we have $S^\dagger=\varepsilon_C\varepsilon_Tu_Cu_T^* =
	\varepsilon_C\varepsilon_T CT$, or $CT= \varepsilon_C\varepsilon_TS^\dagger$.
	Since $TC+CT=0$, we have $S+\varepsilon_C\varepsilon_T S^\dagger=0$, or
	$S^2=-\varepsilon_C\varepsilon_T$.}

@@

Since we are classifying in the solids, it is natural to expect translational
symmetry to present. Also, the addition of translational symmetry not only

produces the same classification\footnote{This is still a mystery to me which this is true. The
	exact reason is related to some isomorphism between certain homotopy groups.
	Interested reader could see Ludwig's lecture
	\href{http://boulderschool.yale.edu/2016/boulder-school-2016-lecture-notes}{here}
	for a clue.}, but also generalize to arbitrary dimensions.
